SPARQL on the Open, Decentralised Web

### 2016-07-24 dissertation/details.tex
intro prelim
details on cartesians

@@

>>>>

This section extends the basic introduction in
Section~\ref{sec:semwebtech} where required to understand the
following discussion.

@@

\subsection{Topics in query evaluation}\label{sec:prelimquery}

As shown in \cite{perez2009semantics} that SPARQL has some highly
complex parts. In particular, nested \sparql{OPTIONAL}s will be
difficult to evaluate. It is possible for a user to either inadvertently
or with malicious intent (in e.g. an denial of service attack) submit
a query that will cause excessive load to the server.

Another problem worthy of special attention is the case where parts of
the query has no connecting variable. Let us consider the
simple example of when two triple patterns do not share a variable:
\begin{verbatim}
?person foaf:name ?name .
?contribution dct:title ?title .
\end{verbatim}
When this is evaluated, it will cause what is known as a ``Cartesian
join'', in which the result will contain \emph{all} the solutions of
the first triple pattern joined with \emph{all} the solutions of the
second triple pattern. This will usually be a large result, and it is
also resource consuming to compute.

In some cases, the users may have written such a query, in which case
the query planner only has the options to execute the query or reject
it entirely. In other cases, the query planner may arrive at such
queries as part of query rewriting, in which case it must ensure that
such plans are not preferred.
